Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2024-12-09 abstract.tex
Draft for committee before defense

This work verified and validated existing multiphysics coupling capabilities in Moltres in two
comparative studies. In both studies, Moltres showed good agreement with other \gls{MSR} simulation tools
involving coupled neutronics and thermal-hydraulics problems. This work also introduces a
turbulence model in Moltres to support future \gls{MSR} analyses involving turbulent delayed
neutron precursor and temperature transport. Lastly, this work introduces a novel hybrid
$S_N$-diffusion method for accurate control rod modeling in time-dependent \gls{MSR} simulations.
The hybrid method combines the strengths of both approaches by generating transport corrections
using the $S_N$ method near control rods. The $S_N$ and neutron diffusion solvers are coupled
through an adaptive boundary coupling algorithm. This algorithm allows the solver to adapt to the
transport correction parameters and preserve smooth neutron flux gradients across the interface.
In 1-D, 2-D, and 3-D $k$-eigenvalue simulations, the hybrid method produced accurate control
rod worth estimates, relative to reference neutron transport solutions and experimental data,
at approximately four times the cost of the neutron diffusion method.
Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
simulations also perfomed well in reproducing expected trends observed in experimental data.
Analysis of its computational performance indicated the possibility of further optimizations
beyond its current implementation. With the hybrid method's spatial resolution and
efficient computational performance, Moltres could enable accurate and cost-effective simulations
of asymmetric transient in \glspl{MSR}.

>>>>

This work verified and validated existing multiphysics coupling capabilities in Moltres in two
comparative studies. In both studies, Moltres showed good agreement with other \gls{MSR} simulation tools
involving coupled neutronics and thermal-hydraulics problems. This work also introduces a
turbulence model in Moltres to support future \gls{MSR} analyses involving turbulent delayed
neutron precursor and temperature transport. Lastly, this work introduces a novel hybrid
$S_N$-diffusion method for accurate control rod modeling in time-dependent \gls{MSR} simulations.
The hybrid method combines the strengths of both approaches by generating transport corrections
using the $S_N$ method near control rods. The $S_N$ and neutron diffusion solvers are coupled
through an adaptive boundary coupling algorithm. This algorithm allows the solver to adapt to the
transport correction parameters and preserve smooth neutron flux gradients across the interface.
In 1-D, 2-D, and 3-D $k$-eigenvalue simulations, the hybrid method produced accurate control
rod worth estimates, relative to reference neutron transport solutions and experimental data,
at approximately four times the computational cost of the neutron diffusion method.
%Demonstrations of the hybrid method for time-dependent rod drop and reactivity insertion
%simulations also performed well in reproducing expected trends observed in experimental data.
A demonstration of the hybrid method for a time-dependent rod drop
simulation also performed well in reproducing expected trends observed in experimental data.
Analysis of its computational performance indicated the possibility of further optimizations
beyond its current implementation. With the hybrid method's spatial resolution and
efficient computational performance, Moltres could enable accurate and cost-effective simulations
of asymmetric transients in \glspl{MSR}.

### 2024-12-09 conclusion.tex
Draft for committee before defense

The 3-D simulations in Chapter \ref{chap:msre} doubled as validation for the 3-D \gls{MSRE} model
against reference \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark study
\cite{fratoni_molten_2020} in
the International Reactor Physics Experiment Evaluation Project (IRPhEP) handbook. OpenMC and
hybrid method estimates of $k_\text{eff}$ of the \gls{MSRE} when the experiment first achieved criticality
$k_\text{eff}$ estimates of the \gls{MSRE} at its initial critcality configuration from OpenMC, the
hybrid method, and the neutron diffusion method in this work showed good agreement with the Serpent
model from the \gls{MSRE} numerical benchmark. All numerical estimates exceeded the experimental
value by 1-2 \% due to possible biases and uncertainties in the nuclear data library for graphite
\cite{fratoni_molten_2020}. Temperature reactivity coefficient values from this work also showed
good agreement with \gls{MSRE} data, within experimental uncertainty, with percentage discrepancies
of about 3 \%. In the subsequent control rod worth study, OpenMC and the hybrid method showed nearly
perfect overlap in the integral rod worth curve throughout the entire length of rod travel. Due to
geometric approximations of the control rods in the numerical models, they overestimated the total
rod worth relative to \gls{MSRE} data by approximately 4-5 \%. The hybrid method significantly
outperforms the neutron diffusion method which overestimated the total worth by 21.1 \%.
When comparing solution times, the hybrid method took approximately four times as long as the
neutron diffusion method.
The 3-D hybrid method simulations exhibited nearly linear scaling in strong scaling tests, which
is promising for future projects involving larger reactor models. Some performance optimizations
may be possible to reduce data transfer times between the $S_N$ and diffusion solvers which took
up about 30 \% of the total solution time due to inter-processor communications.

Lastly, Chapter \ref{chap:transient} demonstrated the hybrid method in time-dependent simulations
based on a \gls{MSRE} zero-power rod drop experiment \cite{prince_zero-power_1968} and a \gls{MSRE}
reactivity insertion experiment during power operation \cite{steffy_experimental_1970}. These
simulations required coupling the hybrid method solver to \gls{DNP} and temperature solvers in
Moltres through nested iteration coupling structures. I set up the rod drop simulation through a
$k$-eigenvalue simulation with the control rod at its initial height before using the simulation
output as the initial condition for the time-dependent rod drop simulation. Moltres reproduced the
expected shape of the integral neutron count curve from \gls{MSRE} experimental data with a maximum
discrepancy of x \% at $t=16$ after the rod drop. Snapshots of the neutron flux spatial
distribution at various timesteps show expected localized flux depressions around the control rod
as it moves through the core.
The reactivity insertion simulation started with the \gls{MSRE} initially at 1 MW power output and
steady salt flow. Following a 24.8 pcm step reactivity insertion, Moltres reproduced the general
shape of the first power peak and the subsequent power oscillations. The power peak magnitude
is y \% higher than the peak in the experimental data. Temperatures in the \gls{MSRE} model rise
and settle at higher values to compensate for the reactivity insertion.

@@

The large 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
with the distributed mesh feature to spread mesh and variable storage across the compute nodes.
Each full-core simulation required at least 40 nodes with 512 GB memory each to run.
This issue may be a stumbling block for using the hybrid method on smaller computing clusters that
have less memory per processor. A custom preconditioner and solver routine in Moltres could help to
reduce memory usage by referencing the same stored Jacobian C++ variable for neutron angular flux
variables in the same neutron energy group and on the same \gls{FEM} quadrature point. Their
Jacobian formulations are identical aside from the level-symmetric ordinate and weight. Another
area for optimization in the hybrid method is the data transfer times between the $S_N$ and
diffusion solvers. Due to mesh distribution, each 3-D simulation spent 30 \% of its solution time
on data transfers between processors. This could be minimized through optimizations in the mesh
distribution and the data transfer caching systems.

>>>>

The 3-D simulations in Chapter \ref{chap:msre} doubled as validation for the 3-D \gls{MSRE} model
against reference \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark study
\cite{fratoni_molten_2020} in
the International Reactor Physics Experiment Evaluation Project (IRPhEP) handbook. OpenMC and
hybrid method estimates of $k_\text{eff}$ of the \gls{MSRE} when the experiment first achieved criticality
$k_\text{eff}$ estimates of the \gls{MSRE} at its initial critcality configuration from OpenMC, the
hybrid method, and the neutron diffusion method in this work showed good agreement with the Serpent
model from the \gls{MSRE} numerical benchmark. All numerical estimates exceeded the experimental
value by 1-2 \% due to possible biases and uncertainties in the nuclear data library for graphite
\cite{fratoni_molten_2020}. Temperature reactivity coefficient values from this work also showed
good agreement with \gls{MSRE} data, within experimental uncertainty, with percentage discrepancies
of about 3 \%. In the subsequent control rod worth study, OpenMC and the hybrid method showed nearly
perfect overlap in the integral rod worth curve throughout the entire length of rod travel. Due to
geometric approximations of the control rods in the numerical models, they overestimated the total
rod worth relative to \gls{MSRE} data by approximately 4-5 \%. The hybrid method significantly
outperforms the neutron diffusion method which overestimated the total worth by 21.1 \%.
When comparing solution times, the hybrid method took approximately four times as long as the
neutron diffusion method.
The 3-D hybrid method simulations exhibited nearly linear scaling in strong scaling tests, which
is promising for future projects involving larger reactor models. Some performance optimizations
may be possible to reduce data transfer times between the $S_N$ and diffusion solvers which took
up about 18 \% of the total solution time due to inter-processor communications.

Lastly, Chapter \ref{chap:transient} demonstrated the hybrid method in a time-dependent simulation
based on a \gls{MSRE} zero-power rod drop experiment \cite{prince_zero-power_1968}.
The simulation required coupling the hybrid method solver to the pre-existing \gls{DNP} solver
in Moltres through a nested iteration coupling setup.
%and a \gls{MSRE}
%reactivity insertion experiment during power operation \cite{steffy_experimental_1970}. These
%simulation required coupling the hybrid method solver to \gls{DNP} and temperature solvers in
%Moltres through nested iteration coupling structures.
I set up the rod drop simulation through a
$k$-eigenvalue simulation with the control rod at its initial height before using the simulation
output as the initial condition for the time-dependent rod drop simulation. Moltres reproduced the
expected prompt and delayed response in the integral neutron count data observed in \gls{MSRE}
experimental data. Convergence issues midway through the simulation had a minor impact on the
solution precision. Overall, this work successfully demonstrated a time-dependent
reactivity-initiated simulation using the hybrid method.
%Snapshots of the neutron flux spatial
%distribution at various timesteps show expected localized flux depressions around the control rod
%as it moves through the core.
%The reactivity insertion simulation started with the \gls{MSRE} initially at 1 MW power output and
%steady salt flow. Following a 24.8 pcm step reactivity insertion, Moltres reproduced the general
%shape of the first power peak and the subsequent power oscillations. The power peak magnitude
%is y \% higher than the peak in the experimental data. Temperatures in the \gls{MSRE} model rise
%and settle at higher values to compensate for the reactivity insertion.

@@

The 3-D simulations in this work uncovered significant memory usage by the $S_N$ solver, even
with the distributed mesh feature to spread mesh and variable storage across the compute nodes.
Each full-core simulation required at least 40 nodes with 512 GB memory each to run.
This issue may be a stumbling block for using the hybrid method on smaller computing clusters that
have less memory per processor. A custom preconditioner and solver routine in Moltres could help to
reduce memory usage by referencing the same stored Jacobian C++ variable for neutron angular flux
variables in the same neutron energy group and on the same \gls{FEM} quadrature point. Their
Jacobian formulations are identical aside from the level-symmetric ordinate and weight. Another
area for optimization in the hybrid method is reducing data transfer times between the $S_N$ and
diffusion solvers. Due to mesh distribution, each 3-D simulation spent 18 \% of its solution time
on data transfers between processors. This could be minimized through optimizations in the mesh
distribution and the data transfer caching systems.
